SeedLing: Building and using a seed corpus \\ for the Human Language Project

### 2014-04-28 docs/computEL-final/cluster.tex
Updated with Liling's new token figures...
changed table -> Table when referencing

Language classification can be carried out in various ways. Two common approaches are genealogical classification, mapping languages onto family trees according to their historical relatedness \cite{swadesh1952,starostin2010}; and typological classification, grouping languages according to linguistic features \cite{georgi2010wals,daume2009}. Both of these approaches require linguistic analysis. By contrast, we use surface features (character n-gram and word unigram frequencies) extracted from SeedLing, and apply an off-the-shelf hierarchical clustering algorithm.\footnote{\url{http://www.scipy.org}} Specifically, each language is represented as a vector of frequencies of character bigrams, character trigrams, and word unigrams. Each of these three components is normalized to unit length.

@@

In table \ref{table:cluster}, we give results of our clustering experiments, comparing the three clustering methods to a random baseline (clustering based on random distances, averaged over 20 runs).  It is worth noting that precision is higher than recall.  This is perhaps to be expected, given that related languages using wildly differing orthographies will appear different.  Nonetheless, our system is reasonably capable of identifying languages which are both related and written similarly.

>>>>

Language classification can be carried out in various ways. Two common approaches are genealogical classification, mapping languages onto family trees according to their historical relatedness \cite{swadesh1952,starostin2010}; and typological classification, grouping languages according to linguistic features \cite{georgi2010wals,daume2009}. Both of these approaches require linguistic analysis. By contrast, we use surface features (character n-gram and word unigram frequencies) extracted from SeedLing, and apply an off-the-shelf hierarchical clustering algorithm.\footnote{\url{http://www.scipy.org}} Specifically, each language is represented as a vector of frequencies of character bigrams, character trigrams, and word unigrams. Each of these three components is normalised to unit length.  Data was taken from ODIN, Omniglot, and the UDHR.

@@

In Table \ref{table:cluster}, we give results of our clustering experiments, comparing the three clustering methods to a random baseline (clustering based on random distances, averaged over 20 runs).  It is worth noting that precision is higher than recall.  This is perhaps to be expected, given that related languages using wildly differing orthographies will appear different.  Nonetheless, our system is reasonably capable of identifying languages which are both related and written similarly.
